---
title: "NIS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NIS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(readHCUP)
```

## The NIS dataset

Working with the National Inpatient Sample (NIS) database can be challenging due to its size (\>7M records per year and 170+ variables) and the limited support for non-proprietary statistical software packages. The goal of `readHCUP` is to make it easier for researchers to focus on their research, not loading data into R. With this in mind, `readHCUP`'s `read_nis()` allows researchers to read datasets with a single function call.

## Working with the NIS dataset

Once you have purchased the NIS dataset from HCUP, you can read the NIS dataset (.ASC file) into R using a single function, `read_nis()`. The following code uses a synthetic NIS 2019 dataset (to avoid publishing real data) to show how to read NIS datasets into R:

```{r}
df <- read_nis("NIS_2019_test_data.ASC", 2019)
```

The path to the dataset and the year it was produced is all we need to read the dataset into R and save it as a tibble.

> NOTE: The current method for reading data using `read_nis()` comes from the `readr` package. In the future, there will be additional support for other methods such as `read.table`

### Corrected datasets

By default, the `read_nis()` automatically returns the corrected version of the data. For example, HCUP released corrections for `PCLASS_ORPROC` in the NIS 2019 and 2020 datasets. Usually, you'd need to download a CSV file with the corrections and then update the values in the dataset. This process can be a bit of a hassle when there are 7M+ records, so the corrections are automatically applied when using `read_nis()`.

-   Note: For the corrections to be applied, `KEY_NIS` and `PCLASS_ORPROC` need to be included in your dataset. If they are not included, `read_nis()` will still return the data, and you will receive a warning that corrections were not applied.

If you don't want the corrections to be automatically applied, use `corrected = FALSE`:

```{r}
# Read dataset the first 10 records of the dataset without corrections.
df <- read_nis("NIS_2019_test_data.ASC", 2019, n_max = 10, corrected = FALSE)
```

#### Supported datasets

The structure of the NIS dataset can change each year, which means `read_nis()` needs to be updated to support each NIS dataset. You can find a list of readHCUP's supported datasets by running the following:

```{r}
supported_datasets
```

-   `data` is the name of the dataset and the year

-   `dataset_file_name` is the file name that was provided by the HCUP Central Distributor

If the dataset is not supported, you will receive an error message:

```{r, error = TRUE}
df_error <- read_nis("nis", 2040)

```

If you're working with a dataset that is not currently supported, please open an [issue in GitHub](https://github.com/jonbry/readHCUP/issues), and we'll work on adding it to the list of supported datasets.

#### Descriptions

The NIS dataset has over 150 variables, which are covered in detail on HCUP's [website](https://hcup-us.ahrq.gov/db/nation/nis/nisdde.jsp). The `descriptions()` function allows you to get a list of all of the variable descriptions:

```{r}
d_list <- descriptions("nis", 2019)
head(d_list)
```
